---
title: "Homework6"
output: html_document
date: "2025-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Load necessary librarues
library(ggplot2)
library(MASS)
z <- read.table("cellular_biology.csv",header=TRUE,sep=",")
str(z)
summary(z)
```
Plot histogram of data
```{r}
# Plot histogram of cell diameters
p1 <- ggplot(data=z, aes(x=Cell_Diameter_um, y=..density..)) +
  geom_histogram(color="grey60", fill="cornsilk", size=0.2) +
  geom_density(linetype="dotted", size=0.75)  # Add empirical density curve
print(p1)
```
The histogram shows the distribution of the data. The empirical density curve helps visualize the spread and shape.

Fit statistical distribution

Fit a normal distribution
```{r}
normPars <- fitdistr(z$Cell_Diameter_um, "normal")
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

# Plot normal probability density
xval <- seq(0, max(z$Cell_Diameter_um), len=length(z$Cell_Diameter_um))
stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", 
                      n = length(z$Cell_Diameter_um), args = list(mean = meanML, sd = sdML))
p1 + stat
```
The red curve represents the normal distribution fit. The mean is slightly biased due to the lack of negative values

Fit an exponential distribution
```{r}
expoPars <- fitdistr(z$Cell_Diameter_um, "exponential")
rateML <- expoPars$estimate["rate"]

# Plot exponential probability density
stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", 
                       n = length(z$Cell_Diameter_um), args = list(rate=rateML))
p1 + stat + stat2
```
The blue curve represents the exponential distribution fit. Exponential distributions model skewed data, but I will compare this to other fits.

Fit a uniform distribution
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", 
                       n = length(z$Cell_Diameter_um), args = list(min=min(z$Cell_Diameter_um), max=max(z$Cell_Diameter_um)))
p1 + stat + stat2 + stat3
```
The green curve represents the uniform distribution, but it does not seem to fit this data set well since it is not uniformly distributed.

Fit a gamma distribution
```{r}
gammaPars <- fitdistr(z$Cell_Diameter_um, "gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

# Plot gamma probability density
stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", 
                       n = length(z$Cell_Diameter_um), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```
The brown curve represents the gamma distribution, which often fits biological data well and appears to be a good fit for this data set.

Fit a beta distribution
```{r}
pSpecial <- ggplot(data=z, aes(x=Cell_Diameter_um/(max(Cell_Diameter_um) + 0.1), y=..density..)) +
  geom_histogram(color="grey60", fill="cornsilk", size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75, linetype="dotted")

betaPars <- fitdistr(x=z$Cell_Diameter_um/max(z$Cell_Diameter_um + 0.1), start=list(shape1=1, shape2=2), "beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", 
                             n = length(z$Cell_Diameter_um), args = list(shape1=shape1ML, shape2=shape2ML))
pSpecial + statSpecial
```
The purple curve represents the beta distribution, but since it assumes the largest point is the true upper bound, it is not the best fit


Determine the best-fitting distribution
Based on the visual comparison, the gamma distribution appears to be the best fit. This makes sense because biological data often follow a gamma distribution due to natural variability.


Simulate a new data set using the best-fitting distribution
```{r}
new_data <- rgamma(n=length(z$Cell_Diameter_um), shape=shapeML, rate=rateML)
df_new <- data.frame(Cell_Diameter_um=new_data)

# Plot histogram of simulated data
p2 <- ggplot(data=df_new, aes(x=Cell_Diameter_um, y=..density..)) +
  geom_histogram(color="grey60", fill="lightblue", size=0.2) +
  geom_density(linetype="dotted", size=0.75)
print(p2)
```
Comparison of original and simulated data
```{r}
p1 + stat4  # Original data with gamma fit
p2  # Simulated data
```

How do the two histogram profiles compare?
  The simulated histogram closely matches the original data. The general shape   and spread are similar. Some minor deviations occur due to randomness in       sampling. 
  
Does the model do a good job of simulating realistic data?
  Yes, the gamma model does a good job because the histogram profiles are very   similar, gamma distributions are often appropriate for biological            
  measurements, and the empirical and modeled density curves align well



  
  
  
