---
title: "Homework4"
output: html_document
date: "2025-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1
Assign values to variables
```{r}
x <- 1.1
a <- 2.2
b <- 3.3
```
Compute each expr4ession and store it in z
Expression 1: 
```{r}
z <- x^(a^b)
print(z)
```

Expression 2:
```{r}
z <- (x^a)^b
print(z)
```

Expression 3:
```{r}
z <- 3*x^3 + 2*x^2 + 1
print(z)
```
#2
Vector 1:
```{r}
vec1 <- c(seq(1,8,1), seq(7,1,-1)
print(vec1)
```

Vector 2:
```{r}
vec2 <- rep(1:5, times 1:5)
print(vec2)
```

Vector 3:
```{r}
vec3 <- rep(5:1, times 1:5)
print(vec3)
```

#3
Generate 2 random uniform numbers (x,y) between 0 and 1
```{r}
set.seed(42) #setting for reproducibility
coords <- runif(2, min=0, max=1)
x <- coords[1]
y <- coords[2]
```
Convert to polar coordinates (r, theta)
```{r}
r <- sqrt(x^2 + y^2) #compute radius
theta <- atan2(y,x) #compute angle in radians
```
Print results
```{r}
cat("Caresian Coordinates: (x,y) =", x, y, "\n")
car("Polar Coordinates: (r, theta) =", r, theta, "\n")
```

#4
Initial queue
```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)
```
a. The serpent arrives and gets in line
```{r}
queue <- c(queue, "serpent")
print(queue)
```
b. The sheep enters the ark (remove first element)
```{r}
queue <- queue[-1]
print(queue)
```

c. The donkey arrives and talks his way to the front of the line
```{r}
queue <- c("donkey", queue)
print(queue)
```
d. The serpent gets impatient and leaves
```{r}
queue <- queue[queue[queue != "serpent"]
print(queue)
```
e. The owel gets board and leaves
```{r}
queue <- queue[queue != "owl"]
print(queue)
```
f. The aphid arrives, and the ant invites him to cut in line (aphid before ant)
```{r}
queue <- append(queue, "aphid", after = match("ant", queue) -1)
print(queue)
```
g. Determine the position of aphid
```{r}
aphid_position <- match("aphid", queue)
cat("The aphid is at position:", aphid_position, "\n"
```
#5
Generate sequence from 1 to 100
```{r}
numbers <- 1:100
```

apply filtering condiiton
```{r}
filtered_numbers <- numbers[numbers %% 2 != 0 & numbers %% 3 != 0 & numbers %% 7 != 0]
```
Print each condition's TRUE/FALSE values
```{r}
print(filtered_numbers)
```

